---
title: "LATENT VARIABLE MEASURE DEVELOPMENT: CFA"
author: "Joanna Moody"
date: 'Last updated: October 16, 2020'
output: html_document
---

```{r setup, results="hide", warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Initialization
rm(list=ls(all=TRUE))
library(knitr)
library(lavaan)
library(semTools)

#Load input data
setwd("/Users/jcmoody/Dropbox (MIT)/07_Latent_attitude_and_AV_perference/")
data <-read.csv("data/individual_new.csv",header=TRUE, sep = ",", stringsAsFactors = FALSE)
data[data == -1] <- NA #-1 indicates missing data

#Note: indicators labeled with "mode_PERC" are "mode_Enjoy"

#Missing data for driving mode: 
#  Individuals who do not have access to a car were not asked these questions;
#  there is no way to estimate a factor score for them because they are missing on
#  all indicators systematically.
#Only estimated for 953 of the 2003 respondents in the sample.

```

## CONVERGENT VALIDITY (continuous)
Treating 7-point scale as continuous (using Maximum Likelihood estimation)

### WALK

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r, linewidth=60}
walk.model0 <- 'Walk =~ WSAFE + WCOMF + WRELY + WEASY + WPERC' 
fit.walk.model0 <- cfa(walk.model0, data=data, missing='fiml')

#Model fit has TLI and RMSEA outside of recommended bounds: 
#  chi-square(N = 2003, df = 5) = 208.247, 
#  CFI = 0.939, TLI = 0.879, 
#  RMSEA = 0.142, SRMR = 0.042
#Factor loadings: lowest standardized loadings is 0.5; all others > 0.6
summary(fit.walk.model0, fit.measures=TRUE, standardized=TRUE)

#Modification indices: suggest that model fit could be greatly improved if 
#  the error terms of the following indicators are correlated: 
#  WSAFE ~~ WRELY are correlated (157.7); WCOMF ~~ WPERC (96.7)
mi.walk.model0 <- modindices(fit.walk.model0)
mi.walk.model0

```

Model 1: Baseline model + correlated error between WSAFE and WRELY

```{r}
walk.model1 <- 'Walk =~ WSAFE + WCOMF + WRELY + WEASY + WPERC
                WSAFE ~~ WRELY' 
fit.walk.model1 <- cfa(walk.model1, data=data, missing='fiml')

#All model fit indices now within recommended bounds: 
#  chi-square(N = 2003, df = 4) = 50.299, 
#  CFI = 0.986, TLI = 0.966, 
#  RMSEA = 0.076, SRMR = 0.020
#Factor loadings: standardized loading for WSAFE dropped to 0.454 (okay, not great)
summary(fit.walk.model1, fit.measures=TRUE, standardized=TRUE)

#Modification indices: MI for WCOMF ~~ WPERC fell to 47.039, so probably not 
#  worth the loss in degrees of freedom to include another correlated error term
mi.walk.model1 <- modindices(fit.walk.model1)
mi.walk.model1

inspect(fit.walk.model1, 'r2')
```


### PUBLIC TRANSIT

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r}
pt.model0 <- 'PT =~ PTSAFE + PTCOMF + PTRELY + PTEASY + PTPERC' 
fit.pt.model0 <- cfa(pt.model0, data=data, missing='fiml')

#Model fit is okay, but should inspect modification indices (slighly high RMSEA: 
#  chi-square(N = 2003, df = 5) = 94.823, 
#  CFI = 0.973, TLI = 0.946, 
#  RMSEA = 0.095, SRMR = 0.028
#Factor loadings: PTSAFE has standardized loading of 0.451, the rest are over 0.65
summary(fit.pt.model0, fit.measures=TRUE, standardized=TRUE)

#Modification indices: 
# Highest MIs are ~45 for PTSAFE ~~ PTEASY and PTSAFE ~~ PTPERC (probably not worth including)
mi.pt.model0 <- modindices(fit.pt.model0)
mi.pt.model0

inspect(fit.pt.model0, 'r2')
```


### RIDEHAILING

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r}
rh.model0 <- 'RH =~ CARSAFE + CARCOMF + CARRELY + CAREASY + CARPERC' 
fit.rh.model0 <- cfa(rh.model0, data=data, missing='fiml')

#Good model fit: 
#  chi-square(N = 2003, df = 5) = 54.086, 
#  CFI = 0.988, TLI = 0.976, 
#  RMSEA = 0.070, SRMR = 0.018
#Factor loadings: All indicators have standardized loadings of > 0.65. Great!
summary(fit.rh.model0, fit.measures=TRUE, standardized=TRUE)

#Very low MIs, so no need for correlated errors (supported by already good model fit) 
mi.rh.model0 <- modindices(fit.rh.model0)
mi.rh.model0 

inspect(fit.rh.model0, 'r2')
```

### DRIVE

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r, warning=FALSE}
drive.model0 <- 'Drive =~ DRSAFE + DRCOMF + DRRELY + DREASY + DRPERC' 
fit.drive.model0 <- cfa(drive.model0, data=data, missing='fiml')

#Very good model fit: 
#  chi-square(N = 953, df = 5) = 10.634,
#  CFI = 0.997, TLI = 0.995, 
#  RMSEA = 0.034, SRMR = 0.011
#Factor loadings: All indicators have standardized loadings of > 0.60!
summary(fit.drive.model0, fit.measures=TRUE, standardized=TRUE)

mi.drive.model0 <- modindices(fit.drive.model0)
mi.drive.model0 
#Low MIs, so no need for correlated errors; can't improve model fit over current levels!

inspect(fit.drive.model0, 'r2')
```


## CONVERGENT VALIDITY (ordinal)
Treating 7-point scale as ordered, categorical data (using WLSMV estimation)

### WALK

Model 1: Baseline model + correlated error between WSAFE and WRELY

```{r}
#walk.model1 <- 'Walk =~ WSAFE + WCOMF + WRELY + WEASY + WPERC
#                WSAFE ~~ WRELY' 
fit.walk.model1.ord <- cfa(walk.model1, data=data, 
                           ordered=c("WSAFE","WCOMF","WRELY","WEASY", "WPERC"),
                           missing="pairwise")

#lavaan using WLSMV estimator: uses diagonally weighted least squares (DWLS) to estimate the model parameters, but uses the full weight matrix to compute robust standard errors, and a mean- and variance-adjusted test stastistic.

#All model fit indices remain within recommended bounds: 
#  chi-square(N = 2003, df = 4) = 28.205, p < .01, scaling correction factor = 0.664
#  CFI = 0.995, TLI = 0.988, 
#  RMSEA = 0.055, SRMR = 0.032
#Factor loadings: standardized loading for all indicators at least 0.5
summary(fit.walk.model1.ord, fit.measures=TRUE, standardized=TRUE)

inspect(fit.walk.model1.ord, 'r2')
```


### PUBLIC TRANSIT

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r}
#pt.model0 <- 'PT =~ PTSAFE + PTCOMF + PTRELY + PTEASY + PTPERC' 
fit.pt.model0.ord <- cfa(pt.model0, data=data, 
                         ordered=c("PTSAFE", "PTCOMF", "PTRELY", "PTEASY", "PTPERC"),
                         missing='pairwise')

#Model fit is okay, but should inspect modification indices (slighly high RMSEA: 
#  chi-square(N = 2003, df = 5) = 33.510, p < .01, scaling correction factor = 0.715
#  CFI = 0.996, TLI = 0.993, 
#  RMSEA = 0.053, SRMR = 0.037
#Factor loadings: All indicators have standardized loading of 0.5 or higher
summary(fit.pt.model0.ord, fit.measures=TRUE, standardized=TRUE)

inspect(fit.pt.model0.ord, 'r2')
```


### RIDEHAILING

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r}
#rh.model0 <- 'RH =~ CARSAFE + CARCOMF + CARRELY + CAREASY + CARPERC' 
fit.rh.model0.ord <- cfa(rh.model0, data=data, 
                         ordered=c("CARSAFE", "CARCOMF", "CARRELY", "CAREASY", "CARPERC"),
                         missing='pairwise')

#Good model fit: 
#  chi-square(N = 2003, df = 5) = 30.188, p < .01, scaling correction factor = 0.674
#  CFI = 0.998, TLI = 0.996, 
#  RMSEA = 0.050, SRMR = 0.023
#Factor loadings: All indicators have standardized loadings of > 0.68. Great!
summary(fit.rh.model0.ord, fit.measures=TRUE, standardized=TRUE)

inspect(fit.rh.model0.ord, 'r2')
```

### DRIVE

Model 0: Baseline model with all 5 indicators, no correlated errors 

```{r, warning=FALSE}
#drive.model0 <- 'Drive =~ DRSAFE + DRCOMF + DRRELY + DREASY + DRPERC' 
fit.drive.model0.ord <- cfa(drive.model0, data=data, 
                            ordered=c("DRSAFE", "DRCOMF", "DRRELY", "DREASY", "DRPERC"),
                            missing='pairwise')

#Very good model fit: 
#  chi-square(N = 953, df = 5) = 3.037, p = 0.69, scaling correction factor = 0.585
#  CFI = 1.00, TLI = 1.00, 
#  RMSEA = 0.000, SRMR = 0.012
#Factor loadings: All indicators have standardized loadings of > 0.65!
summary(fit.drive.model0.ord, fit.measures=TRUE, standardized=TRUE)

inspect(fit.drive.model0.ord, 'r2')
```

## ESTIMATE FACTOR SCORES
Estimate factor scores and append to data records to be included in mode choice models

```{r factscore}
#https://rdrr.io/cran/lavaan/man/lavPredict.html

# With continuous indicators, the possible options for method = are "regression" or "Bartlett"
data$WALK_LV.con <- lavPredict(fit.walk.model1, type = "lv", method = "regression")
data$PT_LV.con <- lavPredict(fit.pt.model0, type = "lv", method = "regression")
data$RH_LV.con <- lavPredict(fit.rh.model0, type = "lv", method = "regression")
data$DRIVE_LV.con <- lavPredict(fit.drive.model0, type = "lv", method = "regression")
#write.csv(data, "data/individual_new_LV.csv")

# With ordinal indicators, the possible options for method = are "EBM" for the Empirical Bayes Modal approach and "ML" for the maximum likelihood approach.
data$WALK_LV.ord <- lavPredict(fit.walk.model1.ord, type = "lv", method = "EBM")
data$PT_LV.ord <- lavPredict(fit.pt.model0.ord, type = "lv", method = "EBM")
data$RH_LV.ord <- lavPredict(fit.rh.model0.ord, type = "lv", method = "EBM")
data$DRIVE_LV.ord <- lavPredict(fit.drive.model0.ord, type = "lv", method = "EBM")

# Calculate correlation between factor scores estimated treating indicators as continuous and factor scores estimated treating indicators as ordinal -- all are at least 0.97
cor(data$WALK_LV.con, data$WALK_LV.ord)
cor(data$PT_LV.con, data$PT_LV.ord)
cor(data$RH_LV.con, data$RH_LV.ord)
cor(data$DRIVE_LV.con, data$DRIVE_LV.ord, use="pairwise.complete.obs")

write.csv(data, "20201016_data-LV.csv")

```

## RELIABILITY

```{r reliability}
#Readily available only for the models treating indicators as continuous (functions do not run with ordinal model fit as input)


#Composite reliability (Omega) and Cronbach's alpha
#  https://rdrr.io/cran/semTools/man/reliability.html 
semTools::reliability(fit.walk.model1)
semTools::reliability(fit.pt.model0)
semTools::reliability(fit.rh.model0)
semTools::reliability(fit.drive.model0)

#Maximal reliability
#  https://rdrr.io/cran/semTools/man/maximalRelia.html
semTools::maximalRelia(fit.walk.model1)
semTools::maximalRelia(fit.pt.model0)
semTools::maximalRelia(fit.rh.model0)
semTools::maximalRelia(fit.drive.model0)

```


## DIVERGENT VALIDITY: Combined Model (continuous)

```{r divergent}
full.model <- ' Walk  =~ WSAFE + WCOMF + WRELY + WEASY + WPERC
                WSAFE ~~ WRELY
                PT =~ PTSAFE + PTCOMF + PTRELY + PTEASY + PTPERC
                RH =~ CARSAFE + CARCOMF + CARRELY + CAREASY + CARPERC
                Drive =~ DRSAFE + DRCOMF + DRRELY + DREASY + DRPERC'

fit <- cfa(full.model, data=data, missing="fiml")
summary(fit, fit.measures=TRUE, standardized=TRUE)
#Model fit is poor: 
#  chi-square(2003, 163) = 3967.0, 
#  CFI = 0.785, TLI = 0.750, 
#  RMSEA = 0.108, SRMR = 0.079
#Currently the model is using FIML to fill in for more than half of the sample missing data on the drive/
#   personal car indicators -- this is extremely suspect!

fit.listwisedelete <- cfa(full.model, data=data) #use default of listwise deletion
summary(fit.listwisedelete, fit.measures=TRUE, standardized=TRUE)
#Model fit is still poor: chi-square(953, 163) = 2381.3, CFI = 0.783, TLI = 0.747, RMSEA = 0.120, SRMR = 0.083
#I suspect there is correlation among the error terms for the same indicators across modes; we can check MIs
mi.fit.listwisedelete <- modindices(fit.listwisedelete, minimum.value=20)
mi.fit.listwisedelete 
#Modification indices:
#  Thankfully, we see very few cross-loadings: RH =~  PTSAFE (64.7), which means our factor structure is clean
#  However, we see a lot of correlated error terms: as expected WSAFE ~~  PTSAFE (192.7), WCOMF ~~  PTCOMF (326.0), 
#     WEASY ~~ PTEASY (131.9)

#But in general, the results above suggest that we should estimate/extract the factor scores for each 
#individual from the separate CFAs rather than from the combined measurement model
```


## DIVERGENT VALIDITY: Combined Model (ordinal)

```{r divergent-ord}
fit.fullmodel.ord <- cfa(full.model, data=data, 
                         ordered=c("WSAFE","WCOMF","WRELY","WEASY", "WPERC",
                                   "PTSAFE", "PTCOMF", "PTRELY", "PTEASY", "PTPERC",
                                   "CARSAFE", "CARCOMF", "CARRELY", "CAREASY", "CARPERC",
                                   "DRSAFE", "DRCOMF", "DRRELY", "DREASY", "DRPERC"),
                         missing='pairwise') #using default of listwise deletion


summary(fit.fullmodel.ord, fit.measures=TRUE, standardized=TRUE)
modindices(fit.fullmodel.ord, minimum.value=100)

#Modification indices:
#We do see some significant cross-loadings and some correlated error terms as we might expect (primarily among the same construct across modes)

#But in general, the results above suggest that we should estimate/extract the factor scores for each individual from the separate CFAs rather than from the combined measurement model (as we have done)
```

